---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    echo: yes
    keep_md: yes
---

## Loading and preprocessing the data
Loading the data from my github repo, and converting date values into POSIXct date format. 
```{r}
temp <- tempfile()
url <- "https://github.com/valpo/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
download.file(url, temp)
actv <- read.csv(unz(temp, "activity.csv"), stringsAsFactors = FALSE)
actv$date <- as.POSIXct(strptime(actv$date,"%Y-%m-%d"))
```

## What is mean total number of steps taken per day?
```{r}
ds <- aggregate(steps ~ date, actv, sum)
hist(ds$steps, main = "Histogram of total number of steps per day", xlab = "total number of steps")
mean.na <- mean(ds$steps)
median.na <- median(ds$steps)
```
The mean and median of the total number of steps taken per day are `r mean.na` and 
`r median.na`.

## What is the average daily activity pattern?
```{r}
ds <- aggregate(steps ~ interval, actv, mean)
plot(ds, type="l")
daily.max <- ds$interval[which.max(ds$steps)]
```
The 5-minute interval that, on average, contains the maximum number of steps is interval 
`r daily.max`. 

## Imputing missing values
The strategy I am using here is:

1. calculate the mean for each interval with na values if we find values for the same 
   interval on other days. 
   
2. set the step value to zero for all other cases. 

```{r}
count.na.before <- sum(is.na(actv))
actvNoNa <- actv
ds <- aggregate(steps ~ interval, actv, mean)
noSteps <- is.na(actvNoNa$steps)
actvNoNa$steps[noSteps] <- ds$steps[ds$interval==actvNoNa$interval[noSteps]]
actvNoNa$steps[is.na(actvNoNa$steps)] <- 0
count.na.after <- sum(is.na(actvNoNa))
```
So we replaced `r count.na.before` occurences of NA and finally got a data.frame containing
`r count.na.after` NA values. 

```{r}
ds <- aggregate(steps ~ interval, actvNoNa, mean)
hist(ds$steps, main = "Histogram of total number of steps per day", xlab = "total number of steps")
mean.nona <- mean(ds$steps)
median.nona <- median(ds$steps)
```
The mean total number of steps taken per day is `r mean.nona`, the corresponding median 
is `r median.nona`. Do these values differ from the values calculated for the databaset before 
replacing NA values?
```{r}
mean.na != mean.nona
median.na != median.nona
```


## Are there differences in activity patterns between weekdays and weekends?
```{r}
library(lattice) 
actvNoNa$weekday <- ifelse(as.integer(strftime(actvNoNa$date, "%u"))<6,"weekday","weekend")
actvNoNa$weekday <- as.factor(actvNoNa$weekday)
ds <- aggregate(steps ~ interval + weekday, actvNoNa, mean)
xyplot(steps~interval|weekday,data = ds, layout=c(1,2), 
       xlab="Interval", ylab="Number of steps", type="l")
```




